Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance

### 2022-08-06 text/Manuscript.Rmd
NIT1_deletion

---
title: Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance
  in zooplankton
author: "Kei Jokura, Nobuo Ueda, Martin Gühmann, Gáspár Jékely"
date: '*Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4
  4QD, UK*'
output:
  html_document:
    theme: lumen
  pdf_document:
    latex_engine: lualatex
    template: default
  word_document: default
geometry: left=0.8in,right=0.8in,top=0.8in,bottom=0.5in
mainfont: Arial
monofont: Times New Roman
documentclass: article
classoption: onecolumn
linestretch: 1.2
fontsize: 10pt
bibliography: references.bib
csl: elife.csl
---

@@

The ability to perceive changes in the environment and adapt behaviour accordingly is essential for the survival of all living organisms. Avoidance from UV radiation, which damages DNA essential for survival, is proposed to be one of the behaviours that existed before the abundant development of Earth's oxygen to block UV radiation in the atmosphere (Pittendrigh). In the ocean, almost all zooplankton have a pronounced rhythmic behaviour called Diel vertical migration (DVM), which consists of upwards towards the surface at dusk and descend back to deeper water before dawn. Avoidance behaviour to escape harmful UV radiation during the day is closely linked to the control of vertical migration in the ocean. In metamorphosis, a life history strategy widely found in metazoans, downward swimming towards the seafloor is known to be associated with larval settlement and the initiation of metamorphosis, a dramatic body change from the larval stage, and UV-avoidance behaviour allows dispersal and the acquisition of different ecological niches. However, little is known about the molecular neural circuit mechanisms in UV avoidance behaviour in many animals, including zooplankton. In neural circuits in general light response, the activation of various receptors by excitatory and inhibitory neurotransmitters such as glutamate, GABA and dopamine is known in various species. In addition to this, it has long been known that in nitric oxide (NO), as a neuromodulatory molecule, it is also important as a modulator of visual activity. NO is a small free-radical molecule with diffusible and changeable characteristics, and thus works as one of the neural signalling molecules. The function of NO in the neural circuitry of vision has been well studied in vertebrates. In the mammalian retina, NO is an essential signalling molecule in all retinal cells. NO produced by nitric oxide synthase (NOS) in amacrine cells, one of the retinal cells, activates soluble guanylate cyclase (sGC) expressed in cone bipolar cells, which produces cGMP and regulates downstream signalling. On the other hand, the role of NO in zooplankton has been reported in many animals, including sponges, cnidarians, molluscs, arthropods, annelids, echinoderms and ascidians, as a molecule that controls settlement and metamorphosis <!---# I think metamorhosis is less relevant for the paper --> initiation. We therefore hypothesise that NO-mediated neural signalling may be involved in the photoreceptor circuits that control UV avoidance behaviour. To test this hypothesis we used the annelid *Platynereis dumerilii*, which conserved vertebrate-type ciliary photoreceptors (cPRCs) [@Arendt2004]. The cPRC expresses c-opsin1, which absorbs UV radiation, is involved in melatonin production and regulates circadian swimming behaviour [@Tosches2014; @tsukamoto2017]. Furthermore, we previously reported abnormal UV avoidance behaviour in c-opsin1 knockout larvae [@Veraszto2018]. The cPRCs express the UV-absorbing c-opsin1 and mediate UV avoidance in the larvae [@Veraszto2018]. In adult worms, UV light and c-opsin1 modulate locomotion and maturation [@Veedin2021]. These suggest that Platynereis larvae are a highly suitable model for elucidating the neural circuit mechanisms of NO-induced UV avoidance behaviour, and the accessibility of genome editing, behavioural assays, calcium imaging and whole-body connectomics could be very useful for study at the molecular level[@ozpolat2021].

>>>>

---
title: Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance
  in zooplankton
author: "Kei Jokura, Nobuo Ueda, Martin Gühmann, Luis Alfonso Yañez-Guerra, Gáspár Jékely"
date: '*Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4
  4QD, UK*'
output:
  pdf_document:
    latex_engine: lualatex
    template: default
  html_document:
    theme: lumen
  word_document: default
geometry: left=0.8in,right=0.8in,top=0.8in,bottom=0.5in
mainfont: Arial
monofont: Times New Roman
documentclass: article
classoption: onecolumn
linestretch: 1.2
fontsize: 10pt
bibliography: references.bib
csl: elife.csl
---

@@

The ability to perceive changes in the environment and adapt behaviour accordingly is essential for the survival of all living organisms. Avoidance from UV radiation, which damages DNA essential for survival, is proposed to be one of the behaviours that existed before the abundant development of Earth's oxygen to block UV radiation in the atmosphere (Pittendrigh). In the ocean, almost all zooplankton have a pronounced rhythmic behaviour called Diel vertical migration (DVM), which consists of upwards towards the surface at dusk and descend back to deeper water before dawn. Avoidance behaviour to escape harmful UV radiation during the day is closely linked to the control of vertical migration in the ocean. In metamorphosis, a life history strategy widely found in metazoans, downward swimming towards the seafloor is known to be associated with larval settlement and the initiation of metamorphosis, a dramatic body change from the larval stage, and UV-avoidance behaviour allows dispersal and the acquisition of different ecological niches. However, little is known about the molecular neural circuit mechanisms in UV avoidance behaviour in many animals, including zooplankton. In neural circuits in general light response, the activation of various receptors by excitatory and inhibitory neurotransmitters such as glutamate, GABA and dopamine is known in various species. In addition to this, it has long been known that in nitric oxide (NO), as a neuromodulatory molecule, it is also important as a modulator of visual activity. NO is a small free-radical molecule with diffusible and changeable characteristics, and thus works as one of the neural signalling molecules. The function of NO in the neural circuitry of vision has been well studied in vertebrates. In the mammalian retina, NO is an essential signalling molecule in all retinal cells. NO produced by nitric oxide synthase (NOS) in amacrine cells, one of the retinal cells, activates soluble guanylate cyclase (sGC) expressed in cone bipolar cells, which produces cGMP and regulates downstream signalling. We therefore hypothesise that NO-mediated neural signalling may be involved in the photoreceptor circuits that control UV avoidance behaviour. To test this hypothesis we used the annelid *Platynereis dumerilii*, which conserved vertebrate-type ciliary photoreceptors (cPRCs) [@Arendt2004]. The cPRC expresses c-opsin1, which absorbs UV radiation, is involved in melatonin production and regulates circadian swimming behaviour [@Tosches2014; @tsukamoto2017]. Furthermore, we previously reported abnormal UV avoidance behaviour in c-opsin1 knockout larvae [@Veraszto2018]. The cPRCs express the UV-absorbing c-opsin1 and mediate UV avoidance in the larvae [@Veraszto2018]. In adult worms, UV light and c-opsin1 modulate locomotion and maturation [@Veedin2021]. These suggest that Platynereis larvae are a highly suitable model for elucidating the neural circuit mechanisms of NO-induced UV avoidance behaviour, and the accessibility of genome editing, behavioural assays, calcium imaging and whole-body connectomics could be very useful for study at the molecular level[@ozpolat2021].
